### 2013-09-13 thesis/src/chapters/chapter-2.tex
Spellings

\subsection{Load-balancing algorithms}
Generally, there are two types of load-balancers: hardware and software based. Due to the dynamic nature of system under consideration, we focus only on the latter as it offers a greater deal of flexibility. Among the most common algorithms we can distinguish \cite{HaProxyDoc}:
\begin{itemize}
 \item \textit{round-robin scheduling} - request are sent to sucessive nodes, according to their weights. This algorithm is fairest when the server's processing time remains equally distributed \cite{HaProxyDoc}
 \item \textit{least connection} - the server with the lowest number of connections receives the connection
 \item \textit{source routing} - source IP address is hashed, the same client IP address always reaches the same server
 \item \textit{URI hashing} - URI that designates resource is hashed and divided by the total weight of the running servers. Such hash designates which server that receives the request. In practice, this algorithm is commonly used with proxy caches and anti-virus proxies in order to maximize the cache hit rate.
 \item \textit{request counting algorithm} - load is distributed the requests among the various workers, ensuring that each gets their configured share of the number of requests
 \item \textit{weighted traffic counting algorithm} - variation of above-mentioned algorithm with a difference that it is focused on bytes rather than number of request
 \item \textit{pending request counting algorithm} - scheduler keeps track of how many requests each worker is assigned at present. A new request is automatically assigned to the worker with the lowest number of active requests
\end{itemize}

@@

While there are many load-balancers available on the market, following are credited to be most popular:
\begin{itemize}
 \item \textbf{HAProxy} \cite{HAProxy} - load-balancer initially written by Willy Tarreau. Noticeably, it's used by OpenShift \cite{OpenShift} to distribute load among gears \cite{OpenShiftScaling}
 \item \textbf{BIG-IP Local Traffic Manager (LTM)} - solution offered by F5 \cite{F5}. Although LTM is a hardware solution, omitted in this section, it also has also its virtualized counterpart.
 \item \textbf{Apache HTTPD} \cite{ApacheHTTPD} - popular HTTP server. When enhanced with additional modules, it can behave like a proxy or load-balancer. Over the time, there were serveral attempts to develop such modules: mod\_jk \cite{ApacheModJk}, mod\_proxy\_balancer \cite{ApacheModProxyBalancer}, to name a few. While the former is purely AJP13 oriented, the latter supports different protocols: HTTP, FTP and AJP13. As a consequnce, only mod\_proxy\_balancer was taken into account during comparision.
\end{itemize}

Table \ref{tab:load-balancer-comparision} presents they key performance features and algorithm used to schedule requests.

@@

\caption{Comparision of load balancers}
\label{tab:load-balancer-comparision}

@@

\\ \hline
OpenVZ &
-- configurable via user beancounters
&
-- configurable via user beancounters 
& 
-- configurable via user beancounters
\\ \hline
\end{tabularx}
\caption{Comparision of hypervisors resizing capabilities}
\label{tab:hypervisors-resizing}
\end{table}

@@

\section{Providers comparision}

Table \ref{tab:cloud-providers-scaling} presents a summary of cloud provider scaling capabilites. Interestingly, all of them are focused solely on horizontal scaling, ignoring advantages offered by a fine-grained approach to scaling that leverage scaling up and application tuning.

@@

\caption{Comparision of cloud providers scaling capabilites}
\label{tab:cloud-providers-scaling}

>>>>

\subsection{Load-balancing algorithms}
Generally, there are two types of load-balancers: hardware and software based. Due to the dynamic nature of system under consideration, we focus only on the latter as it offers a greater deal of flexibility. Among the most common algorithms we can distinguish \cite{HaProxyDoc}:
\begin{itemize}
 \item \textit{round-robin scheduling} - request are sent to successive nodes, according to their weights. This algorithm is fairest when the server's processing time remains equally distributed \cite{HaProxyDoc}
 \item \textit{least connection} - the server with the lowest number of connections receives the connection
 \item \textit{source routing} - source IP address is hashed, the same client IP address always reaches the same server
 \item \textit{URI hashing} - URI that designates resource is hashed and divided by the total weight of the running servers. Such hash designates which server that receives the request. In practice, this algorithm is commonly used with proxy caches and anti-virus proxies in order to maximize the cache hit rate.
 \item \textit{request counting algorithm} - load is distributed the requests among the various workers, ensuring that each gets their configured share of the number of requests
 \item \textit{weighted traffic counting algorithm} - variation of above-mentioned algorithm with a difference that it is focused on bytes rather than number of request
 \item \textit{pending request counting algorithm} - scheduler keeps track of how many requests each worker is assigned at present. A new request is automatically assigned to the worker with the lowest number of active requests
\end{itemize}

@@

While there are many load-balancers available on the market, following are credited to be most popular:
\begin{itemize}
 \item \textbf{HAProxy} \cite{HAProxy} - load-balancer initially written by Willy Tarreau. Noticeably, it's used by OpenShift \cite{OpenShift} to distribute load among gears \cite{OpenShiftScaling}
 \item \textbf{BIG-IP Local Traffic Manager (LTM)} - solution offered by F5 \cite{F5}. Although LTM is a hardware solution, omitted in this section, it also has also its virtualized counterpart.
 \item \textbf{Apache HTTPD} \cite{ApacheHTTPD} - popular HTTP server. When enhanced with additional modules, it can behave like a proxy or load-balancer. Over the time, there were several attempts to develop such modules: mod\_jk \cite{ApacheModJk}, mod\_proxy\_balancer \cite{ApacheModProxyBalancer}, to name a few. While the former is purely AJP13 oriented, the latter supports different protocols: HTTP, FTP and AJP13. As a consequence, only mod\_proxy\_balancer was taken into account during comparision.
\end{itemize}

Table \ref{tab:load-balancer-comparison} presents they key performance features and algorithm used to schedule requests.

@@

\caption{Comparison of load balancers}
\label{tab:load-balancer-comparison}

@@

\\ \hline
OpenVZ &
-- configurable via user beancounters
&
-- configurable via user beancounters 
& 
-- configurable via user beancounters
\\ \hline
\end{tabularx}
\caption{Comparison of hypervisors resizing capabilities}
\label{tab:hypervisors-resizing}
\end{table}

@@

\section{Providers comparison}

Table \ref{tab:cloud-providers-scaling} presents a summary of cloud provider scaling capabilities. Interestingly, all of them are focused solely on horizontal scaling, ignoring advantages offered by a fine-grained approach to scaling that leverage scaling up and application tuning.

@@

\caption{Comparison of cloud providers scaling capabilities}
\label{tab:cloud-providers-scaling}

### 2013-09-13 thesis/src/chapters/chapter-3.tex
Spellings

The remaining of this chapter describes crucial elemenets that compose elasticity controler: policies, data analysis, triggered actions and presents a comparision of cloud providers.

@@

Policy denotes a condition which, when satisfied, triggers an action that is supposed to harness cloud instance in a way that future evalutions of condition will be unsuccessfull. Typically condition itself is accompanied by a minimal and maximal number of node instances, allowing for ensuring minimal QoS and controling maximal costs. Currently, industry leaders supports two main kind of policies \cite{AmazonAutoScaling}:
\begin{itemize}
 \item \textit{expression based} - allows to define how you to scale application in response to changing conditions, which can include factors such as memory, cpu usage, cost or some indirect, calulated metrics
 \item \textit{scheduled} - allows you to scale your application in response to predictable load changes. For example, traffic increases during the weekends and decreases on working days. Hence, we may use that predictable traffic patterns to scale application based on current time.
\end{itemize}

@@

\section{Data analysis}
Having policies defined, their conditions are evaluated against data aquired from sensors. In a simplest case this evaluation can be based on a Threshold Model \cite{LiWoZh05}, which defines a valid range. In cases when given metric violates that condition (i.e. value is either smaller than minimal or bigger than maximal acceptable) corresponding resource is properly adjusted. While trivial in its form, cases of AWS, OpenShift, Carina, OneCloud proves it is useful in a real-world scenarios. Having that said, more sophisticated algorithms also exists:
\begin{itemize}
  \item \textit{integer programming} - autoscaling is reduced to server interger programming problems, which aims to minimize the cost or maximize the computing power with either computing power constraints or
budget constraints \cite{MaLiHu10}
  \item \textit{prediction error correction} - \cite{ShSuGuWi11}
  \item \textit{linearized dynamic control} - linearized correction model is based on control equations \cite{AbShBh02}:

@@

  \item \textit{markov decision process model} - computes optimal reaction to state changes by using obsercation on the system with assumptions about the rate of changes expected in the future \cite{AbWo02}
\end{itemize}

\section{Triggered actions}
Actions that are being triggered by a elasticity controller are focused on application scaling. Problem of scaling an applcication is described in detail in previous chapter.

\section{Providers comparision}

Table \ref{tab:cloud-providers-adaptivity} summaries chapter with approaches to adaptivity taken by different cloud providers. Section 'triggered action' is omited for brevity - it was described extensively in previous chapter.

@@

\caption{Comparision of cloud providers approach to adaptivity}
\label{tab:cloud-providers-adaptivity}

>>>>

The remaining of this chapter describes crucial elements that compose elasticity controller: policies, data analysis, triggered actions and presents a comparison of cloud providers.

@@

Policy denotes a condition which, when satisfied, triggers an action that is supposed to harness cloud instance in a way that future evaluations of condition will be unsuccessful. Typically condition itself is accompanied by a minimal and maximal number of node instances, allowing for ensuring minimal QoS and controlling maximal costs. Currently, industry leaders supports two main kind of policies \cite{AmazonAutoScaling}:
\begin{itemize}
 \item \textit{expression based} - allows to define how you to scale application in response to changing conditions, which can include factors such as memory, cpu usage, cost or some indirect, calulated metrics
 \item \textit{scheduled} - allows you to scale your application in response to predictable load changes. For example, traffic increases during the weekends and decreases on working days. Hence, we may use that predictable traffic patterns to scale application based on current time.
\end{itemize}

@@

\section{Data analysis}
Having policies defined, their conditions are evaluated against data acquired from sensors. In a simplest case this evaluation can be based on a Threshold Model \cite{LiWoZh05}, which defines a valid range. In cases when given metric violates that condition (i.e. value is either smaller than minimal or bigger than maximal acceptable) corresponding resource is properly adjusted. While trivial in its form, cases of AWS, OpenShift, Carina, OneCloud proves it is useful in a real-world scenarios. Having that said, more sophisticated algorithms also exists:
\begin{itemize}
  \item \textit{integer programming} - auto scaling is reduced to server integer programming problems, which aims to minimize the cost or maximize the computing power with either computing power constraints or
budget constraints \cite{MaLiHu10}
  \item \textit{prediction error correction} - \cite{ShSuGuWi11}
  \item \textit{linearised dynamic control} - linearised correction model is based on control equations \cite{AbShBh02}:

@@

  \item \textit{markov decision process model} - computes optimal reaction to state changes by using observation on the system with assumptions about the rate of changes expected in the future \cite{AbWo02}
\end{itemize}

\section{Triggered actions}
Actions that are being triggered by a elasticity controller are focused on application scaling. Problem of scaling an application is described in detail in previous chapter.

\section{Providers comparison}

Table \ref{tab:cloud-providers-adaptivity} summaries chapter with approaches to adaptivity taken by different cloud providers. Section 'triggered action' is omitted for brevity - it was described extensively in previous chapter.

@@

\caption{Comparison of cloud providers approach to adaptivity}
\label{tab:cloud-providers-adaptivity}
